---
title: "hw-04.Rmd"
author: "Gurjot Singh"
date: "09/10/2018"
output: github_document
always_allow_html: yes
---

# Task 1 (Data Reshaping Prompts (and relationship to aggregation))
## Activity #2: Data Reshaping Prompts (and relationship to aggregation)
**Assignment: Make a tibble with one row per year and columns for life expectancy for two or more countries.**

**- Use knitr::kable() to make this table look pretty in your rendered homework.**

**- Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.**

First, lets go ahead and load the packages we will use in this assignment.

```{r}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)
```

Now let's create a new variable, `gapminder_filter` which would filter out the data by year, life expectancies, and by the countries, Canada and Japan.

```{r}
gapminder_filter <- gapminder %>% filter((country) %in% c("Canada", "Japan")) %>% select(country, year, lifeExp)
```

We could use the kable function to make the table more aesthetically appleasing.

```{r}
kable(gapminder_filter[1:24, 1:3], caption = "Countries") %>%
  kable_styling("striped", full_width = F) %>%
  group_rows("Canada", 1, 12) %>%
  group_rows("Japan", 13, 24)
```

Although we used the `kable` function, the graph could be organized in a better way. We can reshape the data by making the country a part of the column section.

```{r}
gapminder_reshape <- gapminder_filter %>% spread(country,lifeExp)

kable(gapminder_reshape) %>%
  kable_styling("striped", full_width = F)
```

Now the list isn't as long as the previous list (`gapminder_filter`).


We could also use ggplot to plot our function.
```{r}
ggplot(gapminder_filter, aes(year,lifeExp)) +
  geom_line(aes(colour=country)) +
  geom_point(aes(colour=country)) +
  ggtitle("Life Expectancy versus Year")
```

It's interesting to note that at the year 1970, there is a point of intersection where Japan's life expectancy becomes higher than Canada's life expectancy.

# Join Prompts (join, merge, look up)
## Activity #2

**Assignment: Create your own cheatsheet patterned after Jenny’s but focused on something you care about more than comics! Inspirational examples:**

**- Pets I have owned + breed + friendly vs. unfriendly + ??. Join to a table of pet breed, including variables for furry vs not furry, mammal true or false, etc.**

**- Movies and studios…**

**- Athletes and teams…**

First, lets create a dataset of different movies. The following varibles will also be recorded: parental advisory, genre, and classification.

```{r}
movies <- "
Name, Parental Advisory,  Genre,  Classification
Shutter Island, Rated R,  Thriller, Fiction
The Pursuit of Happiness, Rated PG-13,  Drama,  Non-fiction
Titanic,  Rated PG-13,  Romance, Non-fiction
Gone Girl,  Rated R,  Thriller, Fiction
Disturbia,  Rated PG-13,  Thriller, Fiction
Game Night, Rated PG-13,  Comedy, Fiction
"
movies <- read_csv(movies, skip = 1)
```

Lets create another dataset which has genre (which is the same column from the above dataset `movies` and could be used to merge the two datasets together) and by a categorical variable which tells us if it is my favourtie genre or not.

```{r}
genre_preference <- "
Genre,  Favourite Genre
Thriller, Yes
Romance,  No
Drama,  Yes
"
genre_preference <- read_csv(genre_preference, skip = 1)
```

Since we have our two datasets (`movies` and `genre_preference`), we could merge them together by using many functions such as:

- `inner_join(x, y)`
- `semi_join(x, y)`
- `left_join(x, y)`
- `anti_join(x, y)`

For ease of understanding I have included the definitions of each function in this assignment:

**inner_join(x, y): Return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join**

```{r}
ijmg <- inner_join(movies, genre_preference)

kable(ijmg) %>%
  kable_styling("striped", full_width = F)
```

It is interesting to note that the movies Game Night is not shown in the merged data as it does not have a corresponding genre with the `genre_preference` dataset.

**semi_join(x, y): Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join**

```{r}
sjmg <- semi_join(movies, genre_preference)

kable(sjmg) %>%
  kable_styling("striped", full_width = F)
```

Using the `semi-join` function we only get variables that are shown in the `movies` dataset.

**left_join(x, y): Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join**

```{r}
ljmg <- left_join(movies, genre_preference)

kable(ljmg) %>%
  kable_styling("striped", full_width = F)
```

The `left_join` function is really great because you get both the `movies` dataset and the `genre_preference` dataset back including the previously ommited Game Night movie which now just has a "NA" under Favourite Genre.

**anti_join(x, y): Return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join**

```{r}
ajmg <- anti_join(movies, genre_preference)

kable(ajmg) %>%
  kable_styling("striped", full_width = F)
```

The `anti_join` function shows the rows from which the previous dataset of `movies` does not have a corresponding set in the `genre_preference` dataset.

# I hope you enjoyed my page!




